Hydra Battles and Cie

### 2021-08-26 doc/part-hydras.tex
Alectryon:  chapter 7 (100%)

\begin{Coqsrc}
 Definition segment (A: Ensemble Ord) :=
  forall alpha beta, In A alpha -> beta < alpha -> In A  beta.

Definition proper_segment (A: Ensemble Ord) :=
  segment A /\ ~ Same_set A ordinal.
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition ordering_function (f : Ord -> Ord)(A B : Ensemble Ord) :=
 segment A /\
 (forall a, In A a -> In B (f a)) /\
 (forall b, B b -> exists a, In A a /\ f a = b) /\
 forall a b, In A a -> In A b -> a < b ->  f a < f b.

Definition ordering_segment (A B : Ensemble Ord) :=
  exists f : Ord -> Ord, ordering_function f A B.
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition the_ordering_segment (B : Ensemble Ord) :=
  the  (fun x => ordering_segment x B).

Definition ord  (B : Ensemble Ord) := 
  some (fun f => ordering_function f (the_ordering_segment B) B).
\end{Coqsrc}

@@

\begin{Coqsrc}
About ordering_function_ex.
\end{Coqsrc}

\begin{Coqanswer}
forall B : Ensemble Ord,
 exists ! S : Ensemble Ord, 
      exists f : Ord -> Ord, ordering_function f S B
\end{Coqanswer}

\begin{Coqanswer}
ordering_function_unicity :
forall (B S1 S2 : Ensemble Ord) (f1 f2 : Ord -> Ord),
ordering_function f1 S B ->
ordering_function f2 S2 B -> 
fun_equiv f1 f2 S1 S2
\end{Coqanswer}

@@

\begin{Coqsrc}
Theorem ordering_le : forall f A B,
    ordering_function f A B ->
    forall alpha, In A alpha -> alpha <= f alpha.

Th_13_5_2 :
forall (A B : Ensemble Ord) (f : Ord -> Ord),
ordering_function f A B -> Closed B -> continuous f A B
\end{Coqsrc}

@@

\index{hydras}{Library Schutte!Functions!plus}
\begin{Coqsrc}
Definition plus alpha := ord  (ge alpha).
Notation "alpha + beta " := (plus alpha beta) : schutte_scope.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma alpha_plus_zero (alpha: Ord): alpha + zero = alpha.
Proof.
 pattern  (plus alpha); apply plus_elim; eauto.
 \end{Coqsrc}

 \begin{Coqanswer}
 1 subgoal (ID 24)

  alpha : Ord
  ============================
  forall f : Ord -> Ord,
  ordering_function f ordinal (ge alpha) -> 
  f zero = alpha
 \end{Coqanswer}

 \begin{Coqsrc}
 (* rest of proof skipped *)
 \end{Coqsrc}

@@

 \begin{Coqsrc}

Lemma zero_plus_alpha (alpha : Ord) : zero + alpha = alpha.

Lemma le_plus_l (alpha beta : Ord) : alpha <= alpha + beta.

Lemma le_plus_r (alpha beta : Ord) :  beta <= alpha + beta.

Lemma plus_mono_r (alpha beta gamma : Ord) : 
    beta < gamma -> alpha + beta < alpha + gamma.

Lemma plus_of_succ (alpha beta : Ord) :
    alpha + (succ beta) = succ (alpha + beta).

Theorem plus_assoc (alpha beta gamma : Ord) :
  alpha + (beta + gamma) = (alpha + beta) + gamma.

Lemma one_plus_omega :  1 + omega = omega.

Lemma finite_plus_infinite (n : nat) (alpha : Ord) :
  omega <= alpha -> n + alpha = alpha.
\end{Coqsrc}

@@

\begin{Coqsrc}
Fixpoint mult_Sn (alpha:Ord)(n:nat){struct n}: Ord :=
 match n with 
            | 0 => alpha
            | S p => mult_Sn  alpha p + alpha
 end.

Definition mult_fin_r alpha n :=
  match n with
      0 => zero
    | S p => mult_Sn alpha p
  end.

Notation "alpha * n" := (mult_fin_r alpha n) : schutte_scope.
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition AP : Ensemble Ord :=
  fun alpha => 
  zero < alpha /\
  (forall beta, beta < alpha ->  beta + alpha = alpha).
\end{Coqsrc}

@@

\index{hydras}{Library Schutte!Functions!phi0}
\begin{Coqsrc}
Definition phi0 := ord AP.

Notation "'omega^'" := phi0 (only parsing) : schutte_scope.
\end{Coqsrc}

@@

\begin{Coqsrc}
Fixpoint omega_tower (i : nat) : Ord :=
  match i with
    0 =>  1
  | S j => phi0 (omega_tower j)
  end.
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition epsilon0 := omega_limit omega_tower.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma AP_one : In AP 1.

Lemma AP_omega : In AP omega.
\end{Coqsrc}

Moreover, $1$ is the least principal ordinal and $\omega$ is the second element of
\texttt{AP}.

\begin{Coqsrc}
Lemma least_AP: least_member  lt AP 1.

Lemma omega_second_AP :
  least_member   lt 
                  (fun alpha => 1 < alpha /\ In AP alpha)
                  omega.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma AP_plus_closed (alpha beta gamma : Ord): 
     In AP alpha -> beta < alpha -> gamma < alpha ->
     beta + gamma < alpha.

Theorem AP_unbounded : Unbounded AP.
\end{Coqsrc}

@@

\begin{Coqsrc} 
Definition Closed (B : Ensemble Ord) : Prop := 
  forall M, Included M B -> Inhabited _ M -> 
                 countable M -> In B (|_| M).
\end{Coqsrc}

\begin{Coqsrc}
Theorem AP_closed : Closed AP.

Lemma AP_o_segment :  the_ordering_segment AP = ordinal.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma AP_phi0 (alpha : Ord) : In AP (phi0 alpha).

Lemma phi0_zero : phi0 zero =  1.

Lemma phi0_mono (alpha beta : Ord) :
  alpha < beta ->  phi0 alpha < phi0 beta.

Lemma phi0_inj (alpha beta : Ord) :
    phi0 alpha = phi0 beta -> alpha = beta.

Lemma phi0_sup : forall (U: Ensemble Ord),
   Inhabited _ U ->   countable U ->  phi0 (|_| U) = |_| (image U phi0).

Lemma is_limit_phi0 (alpha : Ord) :
  zero < alpha ->  is_limit (phi0 alpha).

Lemma omega_eq : omega = phi0 1.

Lemma phi0_le (alpha : Ord) : alpha <= phi0 alpha.
\end{Coqsrc}

Please note that the lemma \texttt{omega\_eq} above, is consistent with the interpretation of the ordering function $\phi_0$ as the exponential of basis $\omega$. Indeed we could have written this lemma with our alternative notation:

\begin{Coqsrc}
 Lemma omega_eq : omega = omega^ 1.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma lt_phi0 (alpha : Ord):
  alpha < epsilon0 -> alpha < phi0 alpha.
\end{Coqsrc}

@@

Moreover,  $\epsilon_0$ is the least ordinal $\alpha$ that verifies the equality 
$\alpha = \omega^\alpha$, in other words the least fixpoint of the function  $\phi_0$.

\begin{Coqsrc}
Theorem epsilon0_lfp : least_fixpoint lt phi0 epsilon0.
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition Cr_fun : forall alpha : Ord,
       (forall beta : Ord, beta < alpha -> Ensemble Ord) ->
        Ensemble Ord 
:= 
   fun (alpha :Ord)
        (Cr : forall beta, 
                beta < alpha -> Ensemble Ord) 
        (x : Ord) => (
       (alpha = zero /\ AP x) \/
       (zero < alpha /\
        forall beta (H:beta < alpha),
          the_ordering_segment (Cr beta H) x /\ ord (Cr  beta H) x = x)).

Definition Cr (alpha : Ord) : Ensemble Ord := 
    (Fix  all_ord_acc (fun (_:Ord) => Ensemble Ord) Cr_fun) alpha.
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition phi (alpha : Ord) : Ord -> Ord 
    :=  ord (Cr alpha).

Definition A (alpha : Ord) : Ensemble Ord :=
  the_ordering_segment (Cr alpha).
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma Cr_zero_AP :  Cr 0 = AP

Lemma epsilon0_Cr1 : In (Cr 1) epsilon0.
\end{Coqsrc}

@@

We prove this result by transfinite induction on $\alpha$ of both properties.

The proof is still quite long, by transfinite induction over $\alpha$.

@@

\begin{Coqsrc}
Section Proof_of_Lemma5.
  Let P (alpha:Ord) := Unbounded (Cr alpha) /\ Closed (Cr alpha).

 Lemma Lemma5 : forall alpha, P alpha.
(* ... *)
 End Proof_of_Lemma5.

Corollary Unbounded_Cr alpha : Unbounded (Cr alpha).
Proof.
  now destruct (Lemma5 alpha).
Qed.

Corollary Closed_Cr alpha : Closed (Cr alpha).
Proof.
  now destruct (Lemma5 alpha).
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
 Definition cnf_t := list Ord.

Fixpoint eval (l : cnf_t) : Ord :=
  match l with nil => zero
              | beta :: l' => phi0 beta + eval l'
  end.

Definition sorted (l: cnf_t) :=
  LocallySorted (fun alpha beta => beta <= alpha) l.

Definition is_cnf_of (alpha : Ord)(l : cnf_t) : Prop :=
  sorted l /\ alpha = eval l.
\end{Coqsrc}

@@

\begin{Coqsrc}
Theorem cnf_exists (alpha : Ord) :
  exists l: cnf_t, is_cnf_of alpha l.
\end{Coqsrc}

@@

\begin{Coqsrc}
 Lemma cnf_unicity : forall l alpha, 
   is_cnf_of alpha l -> 
   forall l',  is_cnf_of alpha l' -> l=l'.
Proof.
 induction l.
 (*  ...  *)

Theorem cnf_exists_unique (alpha:Ord) :
  exists! l: cnf_t, is_cnf_of alpha l.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma cnf_lt_epsilon0 : 
 forall l alpha, 
   is_cnf_of alpha l ->  alpha < epsilon0 ->
   Forall (fun beta =>  beta < alpha) l.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma cnf_of_epsilon0 : is_cnf_of epsilon0 (epsilon0 :: nil).
Proof.
  split.
  - constructor.  
  - simpl;  now rewrite alpha_plus_zero, epsilon0_fxp.
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Fixpoint inject (t:T1) : Ord :=
 match t with 
     | T1.zero => zero
     | T1.ocons a n b =>  AP.phi0 (inject a) * S n + inject b
 end.  
\end{Coqsrc}

This function enjoys good commutation properties with respect to the main operations which
allow us to build Cantor normal form.

\begin{Coqsrc}
Theorem inject_of_zero : inject T1.zero = zero.

Theorem inject_of_finite (n : nat):
  inject (T1.fin n) =  n.

Theorem inject_of_phi0 (alpha : T1):
  inject (phi0 alpha) = AP.phi0 (inject alpha).

Theorem inject_plus (alpha beta : T1): nf alpha -> nf beta ->
  inject (alpha + beta)%t1 = inject alpha + inject beta.

Theorem inject_mult_fin_r (alpha : T1)  :
  nf alpha -> forall n:nat , inject (alpha *  n)%t1 =  inject alpha * n.

Theorem inject_mono (beta gamma : T1) :
  T1.lt  beta gamma -> 
  T1.nf beta -> T1.nf gamma -> 
  inject beta < inject gamma.

Theorem inject_injective (beta gamma : T1) : nf beta -> nf gamma ->
  inject beta = inject gamma -> beta = gamma.
\end{Coqsrc}

@@

\begin{Coqsrc}
Theorem inject_lt_epsilon0 (alpha : T1):
      inject alpha < epsilon0.

Theorem embedding : 
     fun_bijection (nf: Ensemble T1)  (members epsilon0) inject.
 \end{Coqsrc}

>>>>

\input{movies/snippets/Ordering_Functions/segmentDef}

@@

\input{movies/snippets/Ordering_Functions/orderingFunctionDef}

@@

\input{movies/snippets/Ordering_Functions/ordDef}

@@

\input{movies/snippets/Ordering_Functions/orderingFunctionEx}

@@

\input{movies/snippets/Ordering_Functions/orderingLe}
\input{movies/snippets/Ordering_Functions/Th1352}

@@

\index{hydras}{Library Schutte!Functions!plus}

\input{movies/snippets/Addition/additionDef}

@@

As a use-case, let us prove that $0$ is a right neutral element of $+$.

\input{movies/snippets/Addition/alphaPlusZero}

@@

\input{movies/snippets/Addition/bunchOfLemmas}

The following lemmas are not direct applications of \texttt{plus\_elim}.

\input{movies/snippets/Addition/plusAssoc}
\input{movies/snippets/Addition/finitePlusInfinite}

@@

\input{movies/snippets/Addition/multFin}

@@

\input{movies/snippets/AP/APDef}

@@

\index{hydras}{Library Schutte!Functions!phi0}

\input{movies/snippets/AP/phi0Def}

@@

\input{movies/snippets/AP/omegaTower}

@@

\input{movies/snippets/AP/epsilon0Def}

@@

\input{movies/snippets/AP/APOne}

@@

\input{movies/snippets/AP/APPlusClosed}

\input{movies/snippets/AP/APUnbounded}

@@

\input{movies/snippets/Schutte_basics/ClosedDef}
\input{movies/snippets/AP/APClosed}

@@

\input{movies/snippets/AP/APPhi0}

@@

\input{movies/snippets/AP/ltPhi0}

@@

Moreover,  $\epsilon_0$ is the least ordinal $\alpha$ that verifies the equality 
$\alpha = \omega^\alpha$, in other words, the least fixpoint of the function  $\phi_0$.

\input{movies/snippets/AP/epsilon0Lfp}

@@

\input{movies/snippets/Critical/CrDef}

Lets us denote by $\varphi_\alpha$ the ordering function of the set $\textit{Cr}(\alpha)$ and by $A_\alpha$ its ordering segment.

\input{movies/snippets/Critical/phiDef}

@@

@@

\input{movies/snippets/Critical/CrZeroAP}
\input{movies/snippets/Critical/epsilon0Cr1}

@@

We prove this result by a transfinite induction on $\alpha$ of the conjunction of  both properties.

@@

\input{movies/snippets/Critical/Lemma5}

@@

\input{movies/snippets/CNF/Defs}

@@

\input{movies/snippets/CNF/cnfExists}

@@

\input{movies/snippets/CNF/cnfUnicity}

\input{movies/snippets/CNF/cnfExUnique}

@@

\input{movies/snippets/CNF/cnfLtEpsilon0}

@@

\input{movies/snippets/CNF/cnfOfEpsilon0}

@@

This function enjoys good commutation properties with respect to the main operations which
allow us to build Cantor normal forms.

\input{movies/snippets/Correctness_E0/commutationLemmas}
\input{movies/snippets/Correctness_E0/injectPlus}
\input{movies/snippets/Correctness_E0/injectMultFinR}

% \begin{Coqsrc}
% Theorem inject_mono (beta gamma : T1) :
%   T1.lt  beta gamma -> 
%   T1.nf beta -> T1.nf gamma -> 
%   inject beta < inject gamma.

% Theorem inject_injective (beta gamma : T1) : nf beta -> nf gamma ->
%   inject beta = inject gamma -> beta = gamma.
% \end{Coqsrc}

@@

\input{movies/snippets/Correctness_E0/injectLtEpsilon0}
\input{movies/snippets/Correctness_E0/embedding}
